---
title: "Informations"
author: ""
date: ""
output:
   html_document : 
    #css: style.css
   #theme: cosmo  # Style du document
   code_folding: hide #  Cache le code  
   highlight: tango # style de mise en valeur du code
   number_sections: yes # Ajout table des matières
   toc: yes # Table des matière ?
   toc_float: no # table des matière flottante
   toc_depth: 3  # Profondeur table des matière
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R code 

* The application is hosted at the following [URL](https://github.com/proto4426/PissoortThesis/). 
* The **global repository** for the thesis which contains the R package can be found [here](https://github.com/proto4426/PissoortThesis/tree/master/inst/shiny-examples/Bayesian)
* The **specific code**  for this app can be found in this [folder](https://github.com/proto4426/PissoortThesis/tree/master/inst/shiny-examples).

The following package are needed to run the application : 

```{r pack, eval=F}
install.packages(c("shiny", "htmltools", "shinythemes","tidyverse","DT",
                   "gridExtra","grid","mvtnorm","plotly","HDInterval","ggjoy","scales","heatmaply", "coda", "reshape2", "ggcorrplot", "viridis"))
## And of course the package for this thesis ... 
devtools::install_github("proto4426/PissoortThesis", build_vignettes=T)
```

## Model 

* The number of starting values (see diagnostics) influence the total number of iterations ! 
* The seed is set for both the random generation of the starting values (see Algorithm 1 in the text) and the random generation of all the Markov chains. All results can thus be easily retrieved, without having to set two different seeds. 

## Prior

* Yet, these priors are taken to be univariate Normal distributions for each parameters. It would be interesting to incorporate other forms of priors as discussed in 
* The values set by default here and in the thesis are chosen... by default ! **But**, since these are taken to be uninformative (near-flat) priors, relatively small changes won't change the ouput. Verify it by your mouse ! 


## Diagnostics 

The diagnostics presented in Section **4.4** and used in Section **7.4.1** can be used to assess the chains' convergence to their target stationary distribution.


## Posterior Predictive 

Presented in Section **4.6** and used in Section **7.4.3**, the Posterior Predictive Distribution can be conveniently visualized here for the chosen model. 






